### 2020-10-05 paper/introduction.tex

The online reprocessing and refueling system is designed to operate as a long-term reactivity control in the MSRs \cite{ashraf2019modeling,ashraf2019Preliminary}. However, we cannot rely on online reprocessing to adjust reactivity quickly, because loop time (time needed for one particle to make one full circulation through the primary loop) is approximately 20 seconds for MSBR and its analogs. Consequently, we need a prompt reactivity control system for emergencies. The most common and reliable way to control the reactivity is to insert or withdraw control rods made of a material with a large neutron absorption cross section (e.g., boron).
Insertion of control rods introduces negative reactivity into the 
core which helps to compensate the excess reactivity and 
adjust the power level of the core or shut down the reactor in case of 
emergency. Therefore, we should estimate 
the reactivity worth of the control rods.
The reactivity worth of control rods correlates with the interference 
(shadowing effects) between control rod clusters 
\cite{vcerba2017optimization}. 
The control rod worth and its efficiency in compensating excess reactivity is a subject of significant interest since it directly affects the reactor safety \cite{atkinson2019small}. Additionally, the Nuclear Regulatory Commission (NRC) prescribes that one of two independent reactivity control systems should use control rods (Criterion 26) \cite{nuclear1987standard}. According to the demands of nuclear safety regulations, two independent shutdown systems must be adopted in reactor design to ensure the reactor safe operation, to adjust power, and to ensure the shutdown of the reactor anytime independently. All the shutdown systems must be sufficient to shut down the reactor at any time during operation.

>>>>

The online reprocessing and refueling system is designed to operate as a long-term reactivity control in the MSRs \cite{ashraf2019modeling,ashraf2019Preliminary}. However, we cannot rely on online reprocessing to adjust reactivity quickly, because loop time (time needed for one particle to make one full circulation through the primary loop) is approximately 20 seconds for MSBR and its analogs. Consequently, we need a prompt reactivity control system for emergencies. The most common and reliable way to control the reactivity is to insert or withdraw control rods made of a material with a large neutron absorption cross section (e.g., boron).
Insertion of control rods introduces negative reactivity into the 
core which helps to compensate the excess reactivity and 
adjust the power level of the core or shut down the reactor in case of 
emergency. Therefore, we should estimate 
the reactivity worth of the control rods.
The reactivity worth of control rods correlates with the interference 
(shadowing effects) between control rod clusters 
\cite{vcerba2017optimization}. 
The control rod worth and its efficiency in compensating excess reactivity is a subject of significant interest since it directly affects the reactor safety \cite{atkinson2019small}. Additionally, the Nuclear Regulatory Commission (NRC) prescribes that one of two independent reactivity control systems should use control rods (Criterion 26) \cite{nuclear1987standard}. According to the demands of nuclear safety regulations, two independent shutdown systems must ensure the reactor safe operation, adjust power, and ensure the shutdown of the reactor anytime independently. All the shutdown systems must be sufficient to shut down the reactor at any time during operation.

### 2020-10-05 paper/results.tex

The excess reactivity $\rho$$_e$ is calculated by using equation~\ref{Equ:1} at zero burnup (steady-state 
calculation), when all control rods are fully withdrawn. The $\rho_e$ for $^{233}$U, 
reactor-grade Pu, and TRU used as initial fissile material are listed in Table~\ref{tab:excess}.
For $^{233}$U case \cite{ashraf2020Strategies} the initial excess reactivity is about $1180\pm28$ $pcm$, however, the maximum observed excess reactivity is $1344\pm25$ $pcm$ during 60 effective full-power years (EFPY) of operation (see Figure~\ref{fig:keff_25} after $\approx$ 8 EFPY of operation). TRU has a great amount of thermal neutron absorbers, therefore, the SD-TMSR becomes subcritical relatively quickly. Previous studies showed that for promising fueling scenarios (e.g., TRU/$^{232}$Th), large excess reactivity is required for long-term core operation \cite{ashraf2020Strategies,betzler2017assessment,rykhlevskii_fuel_2019}. The proposed reactivity control system must compensate such reactivity at startup and during burnup.

@@

The Shutdown Safety Devices (SSD) clusters are designed mainly for an emergency shutdown, thus it should 
provide the reactor core with sufficient negative reactivity. The 
shutdown margin (SDM) is calculated by equation~\ref{Equ:6}.
Table~\ref{tab:table2} summarizes the shutdown margins for the SD-TMSR core 
initially loaded with $^{233}$U, reactor-grade Pu, and TRU 
elements for different absorbing materials. All absorbing materials provide a sufficient positive SDM for the SD-TMSR core that is initially loaded with $^{233}$U and reactor-grade Pu. We considered the sufficient positive SDM to be 2$\beta$, where $\beta$ is the total delayed fraction. Thus, the sufficient positive SDM is $\sim$1300 $pcm$. The SDMs for the TRU case are negative or slightly positive (in B$_4$C-90 case). Thus, the SSD clusters made of non-B$_4$C-90 materials are ineffectual to shut down the reactor loaded with TRU. Table~\ref{tab:table50} lists the shutdown margin provided by all SSD clusters when a single SSD cluster with maximum worth (i.e., SSD1) is withdrawn. As shown in Table~\ref{tab:table50}, the control rods still meet the shutdown requirement even in the case of SSD1 cluster failure (withdrawal) for the SD-TMSR core that is initially loaded with $^{233}$U and reactor-grade Pu. For the TRU case, in the case of SSD1 cluster failure, substitutional insertion of CSD with SSD clusters will provide a sufficient positive SDM.

@@

We calculated the shutdown margin (SDM) and the amplification factor (A$_{CRi}$) for three different configurations of the CRs in the graphite element for the SD-TMSR initially loaded by TRU (maximum excess reactivity at startup) as demonstrated in Figure~\ref{fig:3_CR}. Results showed that three rods in the graphite element save the 60-degree symmetry of the component; however, the SDM in such a case is $-1058$ $pcm$ and the A$_{CRi}$ is $1.09$. In the case of four rods, the 60-degree symmetry of the graphite element is broken; however, the SDM is about $228$ $pcm$, and the A$_{CRi}$ is $0.97$. Finally, when six rods are distributed evenly in the graphite element, both symmetry and relatively high SDM are obtained ($598$ $pcm$), but the A$_{CRi}$ is about $0.29$, which means high interference between rods due to the shadowing effect in this case. Therefore the 3- and 6-CR configurations are ineffectual due to the negative SDM and the low A$_{CRi}$, respectively. The 4-CRs configuration is adopted in the current study since the SDM is positive, and almost no shadowing effect has been observed.

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{3_CR.png}
	\vspace{-0.5in}
	\caption{$XY$ section of graphite element with a) 3-CRs b) 4-CR c) 6-CRs configurations.} 
	\label{fig:3_CR}
\end{figure}

>>>>

The excess reactivity $\rho$$_e$ is calculated by using equation~\ref{Equ:1} at zero burnup (steady-state 
calculation), when all control rods are fully withdrawn. The $\rho_e$ for $^{233}$U, 
reactor-grade Pu, and TRU used as initial fissile material are listed in Table~\ref{tab:excess}.
For $^{233}$U case \cite{ashraf2020Strategies} the initial excess reactivity is about $1180\pm28$ $pcm$, however, the maximum observed excess reactivity is $1344\pm25$ $pcm$ during 60 effective full-power years (EFPY) of operation (see Figure~\ref{fig:keff_25} after $\approx$ 8 EFPY of operation). The TRU composition includes substantial thermal neutron absorbers, therefore, the SD-TMSR becomes subcritical relatively quickly. Previous studies showed that for promising fueling scenarios (e.g., TRU/$^{232}$Th), large excess reactivity is required for long-term core operation \cite{ashraf2020Strategies,betzler2017assessment,rykhlevskii_fuel_2019}. The proposed reactivity control system must compensate such reactivity at startup and during burnup.

@@

The Shutdown Safety Devices (SSD) clusters are designed mainly for an emergency shutdown, thus it should 
provide the reactor core with sufficient negative reactivity. The 
shutdown margin (SDM) is calculated by equation~\ref{Equ:6}.
Table~\ref{tab:table2} summarizes the shutdown margins for the SD-TMSR core 
initially loaded with $^{233}$U, reactor-grade Pu, and TRU 
elements for different absorbing materials. All absorbing materials provide a sufficient positive SDM for the SD-TMSR core that is initially loaded with $^{233}$U and reactor-grade Pu. We considered the sufficient positive SDM to be 2$\beta$, where $\beta$ is the total fraction of delayed neutron precursors. Thus, the sufficient positive SDM is $\sim$1300 $pcm$. The SDMs for the TRU case are negative or slightly positive (in B$_4$C-90 case). Thus, the SSD clusters made of non-B$_4$C-90 materials are ineffectual to shut down the reactor loaded with TRU. Table~\ref{tab:table50} lists the shutdown margin provided by all SSD clusters when a single SSD cluster with maximum worth (i.e., SSD1) is withdrawn. As shown in Table~\ref{tab:table50}, the control rods still meet the shutdown requirement even in the case of SSD1 cluster failure (withdrawal) for the SD-TMSR core that is initially loaded with $^{233}$U and reactor-grade Pu. For the TRU case, in the case of SSD1 cluster failure, substitutional insertion of CSD with SSD clusters will provide a sufficient positive SDM.

@@

We calculated the shutdown margin (SDM) and the amplification factor (A$_{CRi}$) for three different configurations of the CRs in the graphite element. The SD-TMSR initially loaded by TRU was selected for this analysis due to the maximum excess reactivity at startup. As shown in Figure~\ref{fig:3_CR}, the three rods in the graphite element save the 60-degree symmetry of the component; however, results show that the SDM in such a case is $-1058$ $pcm$ and the A$_{CRi}$ is $1.09$. In the case of four rods, the 60-degree symmetry of the graphite element is broken; however, the SDM is about $228$ $pcm$, and the A$_{CRi}$ is $0.97$. Finally, when six rods are distributed evenly in the graphite element, both symmetry and relatively high SDM are obtained ($598$ $pcm$), but the A$_{CRi}$ is about $0.29$, which means high interference between rods due to the shadowing effect in this case. Therefore the 3- and 6-CRs configurations are ineffectual due to the negative SDM and the low A$_{CRi}$, respectively. The 4-CRs configuration is adopted in the current study since the SDM is positive, and almost no shadowing effect has been observed.

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{3_CR.png}
	\vspace{-0.5in}
	\caption{$XY$ section of graphite element with a) 3-CRs b) 4-CRs c) 6-CRs configurations.} 
	\label{fig:3_CR}
\end{figure}
